--- 
title: "Supplementary materials for paper: Do differences in ceramic incision morphology provide inference to a common toolkit?"
author: 
    - Robert Z. Selden, Jr.^[Heritage Research Center, Stephen F. Austin State University and Cultural Heritage Department, Jean Monnet University, zselden@sfasu.edu]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Morphological analysis of Caddo ceramic incisions"
---

# Preface {-}

>Reproducibility—the ability to recompute results—and replicability—the chances other experimenters will achieve a consistent result—are two foundational characteristics of successful scientific research [@RN20918].

This volume is written in _Markdown_, and all files needed to reproduce it are included in the [GitHub repository](https://github.com/aksel-blaise/incision), as well as the [Open Science Framework (DOI: 10.17605/OSF.IO/SH7WR)](https://osf.io/sh7wr/). The reproducible nature of this undertaking provides a means for others to critically assess and evaluate the various analytical components [@RN20915;@RN20916;@RN20917], which is a necessary requirement for the production of reliable knowledge.

Reproducibility projects in [psychology](https://osf.io/ezcuj/) and [cancer biology](https://www.cos.io/rpcb) are impacting current research practices across all domains. Examples of reproducible research are becoming more abundant in archaeology [@RN20804;@RN21009;@RN21001], and the next generation of archaeologists are learning those tools and methods needed to reproduce and/or replicate research results [@RN21007]. Reproducible and replicable research work flows are often employed at the highest levels of humanities-based inquiries to mitigate concern or doubt regarding proper execution, and is of particular import should the results have—explicitly or implicitly—a major impact on scientific progress [@RN21008].

Components of the undertaking presented in this volume are novel, and this basic research endeavour is rooted in archaeological epistemology. Sharing data is not enough, and if investigators share only data absent their analysis code, that places a substantive burden on those who may seek to build upon or replicate their work in the future. In this instance, the choice to pursue a replicable work flow occurs based on the simple fact that this work is foundational, and meant to begin a meaningful discussion related to the similarities and differences between tools used to apply—whether incised, engraved, or etched—the many decorative motifs that adorn ceramic artefacts.

As an exploratory research endeavour, this study was not [preregistered](https://www.cos.io/prereg).

## Acknowledgments

I extend my gratitude to the Caddo Tribe of Oklahoma and the National Forests and Grasslands in Texas for the requisite permissions and access needed to generate the scans of the Caddo incisions, and to the National Center for Preservation Technology and Training for access to the optical profilometer. 

This volume enlists a variety of tools from the [Open Review Toolkit](https://www.openreviewtoolkit.org/), as well as the code provided for the `bookdown` package. I extend my gratitude to all who contribute comments and constructive criticisms throughout the development and maturation of this project. This document will remain in open review until the article is published.

## Funding

Components of the analytical workflow were developed and funded by a Preservation Technology and Training grant (P14AP00138) to RZS from the National Center for Preservation Technology and Training, as well as grants to RZS from the Caddo Nation of Oklahoma, National Forests and Grasslands in Texas (15-PA-11081300-033) and the United States Forest Service (20-PA-11081300-074).

## Colophon

This version of the analysis was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# what R packages and versions were used?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

Current Git commit details are:

```{r}
# where can I find this commit? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

