--- 
title: "Toward a theory of the mark: A pilot study of ceramic incision morphology"
author: 
    - Robert Z. Selden, Jr.^[Heritage Research Center, Stephen F. Austin State University; Department of Biology, Stephen F. Austin State University; and Cultural Heritage Department, Jean Monnet University, zselden@sfasu.edu]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Morphological analysis of ceramic incisions"
---

# Preface {-}

>Each [Caddo] clan had its own shape to make its pottery. One clan never thought of making anything the same pattern of another clan. **_You could tell who made the pottery by the shape_** [@RN9357x, 395].

## Basis of inquiry

Decorative motifs that adorn prehistoric ceramic vessels are employed by archaeologists as temporally, spatially, and culturally diagnostic traits, and are often used to characterize diagnostic archaeological types. Among the distinguishing marks that adorn prehistoric ceramics, incisions capture the most complete morphological record of the tools used to apply them. This is due primarily to the fact that incised marks were applied while the clay matrix included a higher water content, providing a malleable plastic surface where deep, high-relief marks could be applied easily to vessel surfaces. The deepest incision profiles provide the most complete cross section of those tools used to apply them. A macro analysis of incised Caddo ceramics from sites in the National Forests and Grasslands in Texas suggests that two discrete shapes and three distinct tool profiles were used in the application of incisions by Caddo potters. To evaluate the extent to which incision profiles differ, incisions made on experimental and archaeological samples were scanned using an optical profilometer, and incision morphology was analysed using the tools of geometric morphometrics. In experimental and archaeological samples, three discrete---and significantly different---incision profiles were identified. Results raise questions regarding whether a common tool or toolkit may have been used to apply incisions that articulate with a specific motif or group of motifs

## Keywords {-}
American Southeast, Caddo, computational archaeology, digital humanities, museum studies, non-Western art history, STEM, STEAM

## Acknowledgments {-}

I extend my gratitude to the Caddo Nation of Oklahoma and the National Forests and Grasslands in Texas for the requisite permissions and access needed to generate the scans of the Caddo incisions, and to the National Center for Preservation Technology and Training for access to the optical profilometer. 

This volume enlists a variety of tools from the [Open Review Toolkit](https://www.openreviewtoolkit.org/) and the `bookdown` package. I extend my gratitude to all who contribute comments and constructive criticisms throughout the development and maturation of this project. This document will remain in open review until the article is published.

## Funding {-}

Components of the analytical workflow were developed and funded by a Preservation Technology and Training grant (P14AP00138) to RZS from the National Center for Preservation Technology and Training, as well as grants to RZS from the Caddo Nation of Oklahoma, National Forests and Grasslands in Texas (15-PA-11081300-033) and the United States Forest Service (20-PA-11081300-074). Additional support for this project was provided by the Heritage Research Center at Stephen F. Austin State University.

## Data management {-}

Reproducibility—the ability to recompute results—and replicability—the chances other experimenters will achieve a consistent result—are two foundational characteristics of successful scientific research [@RN20918]. To that end, this volume is written in _Markdown_, and all files needed to reproduce it are included in the [GitHub repository](https://github.com/aksel-blaise/incision), and digitally curated on the [Open Science Framework (DOI: 10.17605/OSF.IO/SH7WR)](https://osf.io/sh7wr/). The reproducible nature of this undertaking provides a means for others to critically assess and evaluate the various analytical components [@RN20915;@RN20916;@RN20917], which is necessary for the production of reliable knowledge.

Reproducibility projects in [psychology](https://osf.io/ezcuj/) and [cancer biology](https://www.cos.io/rpcb) are impacting current research practices across all domains. Examples of reproducible research are becoming more abundant in archaeology [@RN20804;@RN21009;@RN21001], and the next generation of archaeologists are learning those tools and methods needed to reproduce and/or replicate research results [@RN21007]. Reproducible and replicable research work flows are often employed at the highest levels of humanities-based inquiries to mitigate concern or doubt regarding proper execution, and is of particular import should the results have—explicitly or implicitly—a major impact on scientific progress [@RN21008].

## Colophon {-}

This version of the analysis was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# what R packages and versions were used?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

Current Git commit details are:

```{r}
# where can I find this commit? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
